---
title: 'Suggested solutions for Homework 0'

date: "9/6/2022"
output: html_document
---


## 1. Understanding the Data
```{r}
library(readr)

# Loading the dataset
covid_df <- read_csv("covid19.csv")
```

```{r}
# Display the dimension of the data: 
dim(covid_df)

# Store the column names in a variable
vector_cols <- colnames(covid_df)

# Display the variable vector_cols
vector_cols

# Explore the first 5 rows
head(covid_df)

# Showing a global view of the dataset.
library(tibble)

glimpse(covid_df)

```

The dataset contains `14` columns and `10,903` rows. This database provides information on the numbers (per day and cumulatively) of COVID-19 positive cases, deaths, tests performed and hospitalizations for each country through the column's names store in the variable `vector_cols`. 

1. This variable contains a character vector.

2. The use of the function `glimpse()` is the very first operation to do because we don't only learn about the dimensions of the database but also about the names of the first columns and their types and content. We don't have to rely on `glimpse()`, but at least need to show some summaries.

## 2. Isolating the Rows 

After reviewing the question, it seems answerable. In fact, it contains too much information. 

- Selecting only the rows related to `"All States"` and removing the `Province_State`.

```{r}
library(dplyr)

# Filter the "All States" Province states and remove the `Province_State` column
covid_df_all_states <- covid_df %>% 
  filter(Province_State == "All States") %>% 
  select(-Province_State)

```
- We can remove `Province_State` without loosing information because after the filtering step this column only contains the value `"All States"`.

## 3. Isolating the Columns
- Creating a dataset for the daily columns from `covid_df_all_states` dataframe

Let's recall the description of the dataset's columns.

1. `Date`: Date
2. `Continent_Name`: Continent names
3. `Two_Letter_Country_Code`: Country codes
4. `Country_Region`: Country names
5. `Province_State`: States/province names; value is `All States` when state/provincial level data is not available
6. `positive`: Cumulative number of positive cases reported.
7. `active`: Number of actively cases on that **day**.
8. `hospitalized`: Cumulative number of hospitalized cases reported.
9. `hospitalizedCurr`: Number of actively hospitalized cases on that **day**.
10. `recovered`: Cumulative number of recovered cases reported.
11. `death`: Cumulative number of deaths reported.
12. `total_tested`: Cumulative number of tests conducted.
13. `daily_tested`: Number of tests conducted on the **day**; if daily data is unavailable, daily tested is averaged across number of days in between.
14. `daily_positive`: Number of positive cases reported on the **day**; if daily data is unavailable, daily positive is averaged across number of days in.


```{r}
# Selecting the columns with cumulative numbers
covid_df_all_states_daily <- covid_df_all_states %>% 
  select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

head(covid_df_all_states_daily)
```


## 4. Extracting the Top Ten countries in the number of tested cases

### Summarize the data based on the `Country_Region` column.
```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested = sum(daily_tested), 
            positive = sum(daily_positive),
            active = sum(active),
            hospitalized = sum(hospitalizedCurr)) %>% 
  arrange(desc(tested)) #this is equivalent to `arrange(-tested)`

covid_df_all_states_daily_sum


```

## 5. Identifying the Highest Positive Against Tested Cases

##
```{r}
covid_df_positive_per_test <- covid_df_all_states_daily_sum %>%
      mutate(
        postive_per_test = positive / tested
      ) %>%
      arrange(desc(postive_per_test))
head(covid_df_positive_per_test)
tail(covid_df_positive_per_test)
```

## 6. Now, what would you do?