---
title: "Homework4"
author: "Yeabin Moon"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 4

## Part 1

In this project, we'll work with data from the CIA World Factbook, a compendium of statistics about all of the countries on Earth. The Factbook contains demographic information like the following:

- population — the global population.
- population_growth — the annual population growth rate, as a percentage.
- area — the total land and water area.

### 1. Write a query to return information on the tables in the database.
```
Write a query
```
```{r echo = F}
print('show your result here')
```

You will see three tables in the database. See 'facts' table first.
Here are the descriptions for some of the columns:
```
name — the name of the country.
area— the country's total area (both land and water).
area_land — the country's land area in square kilometers.
area_water — the country's waterarea in square kilometers.
population — the country's population.
population_growth— the country's population growth as a percentage.
birth_rate — the country's birth rate, or the number of births per year per 1,000 people.
death_rate — the country's death rate, or the number of death per year per 1,000 people.
```
Let's start by calculating some summary statistics and look for any outlier countries.

### 2. Write a single query that returns the following:
```
Write a query
```
```{r echo = F}
print('show your result here')
```
- Minimum population
- Maximum population
- Minimum population growth
- Maximum population growth

We see a few interesting things in the summary statistics from the previous question.

- There's a country with a population of 0
- There's a country with a population of 7256490011 (or more than 7.2 billion people)

Let's use subqueries to zoom in on just these countries without using the specific values.

### 3. Write a query that returns the countries with the minimum population.
```
Write a query
```
```{r echo = F}
print('show your result here')
```

### 4. Write a query that returns the countries with the maximum population.
```
Write a query
```
```{r echo = F}
print('show your result here')
```


It seems like the table contains a row for the whole world, which explains the population of over 7.2 billion. It also seems like the table contains a row for Antarctica, which explains the population of 0. Now that we know this, we should recalculate the summary statistics we calculated earlier — this time excluding the row for the whole world.

### 5. Recompute the summary statistics you found earlier while excluding the row for the whole world. Include the following:

- Minimum population
- Maximum population
- Minimum population growth
- Maximum population growth

### 6. In a different code cell, calculate the average value for the following columns:

- population
- area
```
Write a query
```
```{r echo = F}
print('show your result here')
```

To finish, we'll build on the query we wrote for the previous screen to find countries that are densely populated.

### 7. Write a query that finds all countries meeting both of the following criteria:

- The population is above average.
- The area is below average.

## Part 2
Review the lecture 12 before proceed. we'll work with data 'chinook.db'.

In some cases, there can be a relation between two columns within the same table. Within the employee table, the reports_to column has a relation to the employee_id column within the same table.

The reports_to column identifies the employee ID of each employee's supervisor. Since employee_id is the primary key for this table, we would interpret reports_to as a reference to another row in the table. If we wanted to create a report of each employee and their supervisor's name, we would need some way of joining a table to itself. This type of join is called a <b>recursive join</b>.

Syntactically, a recursive join will resemble any of the other joins we've learned before. The only difference is that we're joining a table to itself. We'll use aliases to distinguish between the two sides of the join. On the left side, e1, we focus on the reports_to column since it refers to their supervisor's ID. For the right side, the focus is on the employee_id since we are trying to match the supervisors by their own employee ID. Here's a simple example of a recursive join in action:

```
See this qurey produces:
SELECT
    e1.employee_id,
    e2.employee_id supervisor_id
FROM employee e1
INNER JOIN employee e2 on e1.reports_to = e2.employee_id
LIMIT 4;
```

We'll add some extra columns to make our report more meaningful. Instead of using IDs, we'll create a new column with the full name of the employee. To combine columns, we can use the <b>concatenate operator</b>: ||. We'll concatenate the first_name and last_name columns to get the desired full name.

You can explore how the concatenate operator works by writing a simple query:
```
SELECT ("this" || "is" || "my" || "string");
```

Here's a few different examples of using concatenation, this time with columns from the album table:
```
SELECT
    album_id,
    artist_id,
    "album id is" || album_id col_1,
    "artist id is" || artist_id col2,
    album_id || artist_id col3
FROM album LIMIT 3;
```

One nice perk you may have noticed is that the SQL engine handles converting different types where needed.

Let's use what we've learned about recursive joins and the concatenation operator to create a report about employees and their supervisors.

### 8.
1. Write a query that returns information about each employee and their supervisor. Your query should return the following columns, in order:

  - employee_name - containing the first_name and last_name columns separated by a space, e.g. Luke Skywalker.
  - employee_title - the title of that employee.
  - supervisor_name - the first and last name of the person the employee reports to, in the same format as employee_name.
  - supervisor_title - the title of the person the employee reports to.
2. The report should include employees even if they do not report to another employee.
3. The report should be sorted alphabetically by the employee_name column.


## Part 3
See the slide 20 and 21 in lecture 12. We see that as a sample size grows, the average gets closer to the average of the population (whole) data. Why does a sample size matters here? 