---
title: "Homework3"
author: "Yeabin Moon"
date: "2022-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use gapminder dataset. 
```{r}
library(gapminder)
library(tidyverse)
```

### 1. Consider the following: Explain what relationship the figure represents.
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = pop,
                          y = lifeExp)) 
p + geom_point() +
    geom_smooth(color = "orange") 
```


### 2. What does the orange line represent?

### 3. You can use a range of methods in geom_smooth(). What is the default method? Try to plot four figures using different methods in geom_smooth(). 

### 4. What is the grey area around the trend line? Modify the code in question 1 and plot the figure without it.

### 5. We learned in class the scale mapping, such as scale_x_log10(). Search online and list two more commands (functions) to transform the value of x.

### 6. Explain the difference between the two code blocks
```{r}
# code 1
p <-  ggplot(data = gapminder,
             mapping = aes(x = gdpPercap,
                            y = lifeExp,
                            color = year))
p + geom_point() +
    geom_smooth(method = "loess", se = F) + scale_x_log10()
```
```{r}
# code 2
p <-  ggplot(data = gapminder,
             mapping = aes(x = gdpPercap,
                            y = lifeExp,
                            color = factor(year)))
p + geom_point() +
    geom_smooth(method = "loess", se = F) + scale_x_log10()

```

### 7. Explain the difference between facet_wrap() and facet_grid(). Show the examples. You cannot use the same data in the lecture example.