---
title: "Lecture18"
author: "Yeabin Moon"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
setwd("~/Documents/ibs_course/BUS211/data") # Set your path
dc_listings <- read_csv('tidy_dc_airbnb.csv')
```

## 1. Introducing The caret Library

Let's do

1. Load in the caret library.
2. Use the createDataPartition() function to perform the train/test split on the dc_listings data. More specifically:
    - allocate 70% of the data to the training set,
    - make sure to return a vector instead of a list, and
    - assign the output to the name train_indices
3. Use the train_indices variable to actually split the dc_listings data. Assign the training set to the variable train_listings and the test set to test_listings

```{r}
set.seed(1)
library(caret)

train_indices <- createDataPartition(y = dc_listings[["tidy_price"]],
                                     p = 0.7,
                                     list = FALSE)
train_listings <- dc_listings[train_indices,]
nrow(train_listings)
test_listings <- dc_listings[-train_indices,]
nrow(test_listings)
nrow(train_listings) / (nrow(train_listings)+ nrow(test_listings))
```

## 2. Training The Algorithm

```{r}
# 
train_control <- trainControl(method = "none")
```

Create a trained k-nearest neighbors model using the train() function. Assign this trained model to the knn_model variable.
    - Predict tidy_price from the accommodates and maximum_nights columns.
    
```{r}
knn_model <- train(tidy_price ~ accommodates + maximum_nights, 
                   data = train_listings, 
                   method = "knn", 
                   trControl = train_control)
```

## 3. Create Predictions On The Test Data

Create predictions for the test_listings. Assign your predictions to the variable test_predictions.
    - Use both the accommodates and maximum_nights as your features in this algorithm.
```{r}
test_predictions <- predict(knn_model, newdata = test_listings)
```

## 4. Evaluating Predictions

Create a new column in test_listings called error that captures how much the model predictions differ from the actual listing prices.

```{r}
test_listings <- test_listings %>%
  mutate(
    error = tidy_price - test_predictions
  )

```

## 5. Summarizing Errors Into a Single Metric
1. First, calculate the squared error based on the original error column that you created on the last screen. Give this column the name squared_error.
2. Using this new squared_error column, calculate the RMSE. Assign this to the variable rmse.
```{r}
test_listings <- test_listings %>%
  mutate(
    squared_error = error^2
  )

rmse <- sqrt(mean(test_listings$squared_error))
rmse
```