---
title: "Homework 1"
author: "Yeabin Moon"
date: "2022-09-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Learning R environment

To program with R, it's important to understand how values and data types work. Let's start with literal values. 

A literal value is part of a statement or expression used exactly as intended for interpretation. For instance, the numbers 125 and 3 are literal values because they literally represent the numbers 125 and 3.

Every literal value has a type. In R, there are five basic data types: Logical, Integer, Numeric, Character, and Complex. First, we'll focus on the Logical data type.

With the Logical data type, there are two literals, "true" and "false". In R, we write TRUE and FALSE — or simply T and F. All the letters are in UPPERCASE, because R is case-sensitive it differentiates between uppercase and lowercase.

In programming languages, the logical literals represent two values: 1 (for true) and 0 (for false). Similarly, R also counts the TRUE literal as 1 and the FALSE literal as 0.

Write TRUE or FALSE in the code block to indicate if the following statement is correct or incorrect.

1. F or FALSE represents the "false" literal in R.
```{r}
print('write your answer')

```

There are two types of numbers in R: integers (Integer) and decimal/real (Numeric) numbers. To force R to recognize numbers as integer literals (without the decimal part), the letter L must accompany the number for example 50L.

2. Enter 100 so that R considers it an integer literal. Check your answer using class function
```{r}
print('write your answer')
```

R interprets a number as a Character if we enclose it in quotation marks or apostrophes (e.g., "3.14"). Play with your own numbers/characters in R

## Data type transofrmation rules
Operations between values of the same data type yield that same data type. For example, 2L+2L provides an Integer data type.

Operations between values of different data types will result in the highest data type. From highest to lowest, the data types rank in this order: Numeric, Integer, and Logical.

3. What is the type of the following expression: 2 * 2 * 2?
```{r}
print('write your answer')
```

4. What is the type of the following expression: 2 * 2 * 2L?
```{r}
print('write your answer')
```

## Logical Expressions in R

We use relational operators to compare between values. So, comparison operations return logical values TRUE or FALSE.

5. List all the relational operators available in R:
```
Write your answer in words
```


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
