---
title: "Lecture5"
author: "Yeabin Moon"
date: "2022-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction to Introduction to R

Here are a few topics we will cover in this lesson:

- The basics of R syntax
- Introduction to data types in R
- How to perform calculations using comparison operators


## 1.1 Console and script
In the script or code editor, you can write the instructions. 
```{r}
1223189283+2
```
Run this commend either in the script or console.

What is [1] in front of the result? Try this:
```{r}
1:100
```

## 1.2 Data type
In R, there are five basic data types: Logical, Integer, Numeric, Character, and Complex.

Logical data type consists of TRUE and FALSE
```{r}
TRUE == T
TRUE == 1
```
```
Try: TRUE == True
```
R is case-sensitive!

There are two types of numbers in R: Integer and Numeric. You can think that whole numbers and other numbers. Note that R recognizes any number as Numeric type. We need to put "L" after whole number to make R understand it as Integer
```{r}
100L
99999L
```

To represent character, we need either quotation marks " or apsstrophes '
```{r}
"Hi"
"Have you put enough effort?"
"'FALSE' is not logical literal any more"
"'1' is a character"
```

# 2. Arithmetic
You know how to do arithmetic. I want to show something more.

## 2.1 R knows the operator priority rules
```{r}
1+2*3
(1+2)*3
```

## 2.2 Data type
Recall that there are two types of number. Operations among values of the same data type yield the same data type
```{r}
result1 <- 123L + 123L * 2L
print(result1)
class(result1)
result2 <- 123 + 123 * 2
print(result2)
class(result2)
```
Note that data types has a descending order: Numeric, Integer, Logical. Operations among different data type yield the highest data type 
```{r}
result3 <- 123L + 123L * 2
print(result3)
class(result3)
```

## 2.3 Variables
We save a result into a variable in many purposes. The variable is nothing but a position in your machine in which your computer saves the result. Each programming language has a rule on variable naming. Here is the R:

1. Name consists of letters, numbers, ., _ (no comma)

2. The first letter should be a letter or .
  - You cannot start with a number
  - You can start with . but then number cannot be followed
```
Try: .a<-2
Try: .2<-2
Try: 1998aa<-2
```
3. Special cahracters are not allowed (spaces, operators,", parentheses)

4. Don't use a keyword already used in R

```{r}
here_are_some_examples_including_spaces_variables <- 10
```

## 2.4 Relational Operators
Relational operators are used to compare between values. So it returns logical values either TRUE or FALSE.

- Less than (<)
- Greater than (>)
- Less than or equal to (<=)
- Greater than or equal to (>=)
- Equal to (==)
- Not equal to (!=)

```{r}
223L < 4000L
!TRUE == FALSE
```

Consider this:
```
11 - 23123 > 11 + 111111111
```
Note that all arithmetic expression are evaluated before logical expressions.

## 2.5 Logical operators
- Logical NOT (!)
- Logical AND (&&)
- Logical OR (||)
Consider:
```
TRUE && FALSE
TRUE || FALSE
!TRUE
```
<b>Try to think about how to code $-1< x< 2$. </b>

# 3. Data basics
R contains a lot of useful toolbox to perform data work. These tools are collection of functions which are stored in packages. 

To install an R package, we use the function <b>install.packages()</b>.

