---
title: "Lecture6"
author: "Yeabin Moon"
date: "2022-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Vector (continued)

## 1.1 Vector operations
Consider the following vectors. Put the comments.
```{r}
exam_grades    <- c(92, 90, 84, 95, 77, 92, 85)
homework_grades<- c(87, 81, 95, 86, 85, 90, 88)
project_grades <- c(85, 92, 79, 93, 90, 91, 95)

class_scores_sum <- exam_grades + homework_grades + project_grades # put your comment here
class_scores_avg <- class_scores_sum / 3                           # put your comment here
```

For the same length vector, it is not hard to expect that element-wise operations are performed.

## 1.1 Recycling rule
It might look very confusing:
```{r}
vec1 <- c(1,2,3,4,5)
vec2 <- c(10,10)
vec3 <- vec1 + vec2
print(vec3)

vec4 <- c(10,100)
vec5 <- vec1+vec4
print(vec5)
```

### Exercise 1
1. Calculate the sum score for each class by adding the exam_grades, homework_grades, and project_grades vectors.
2. Calculate the average scores for each class by dividing the class_scores_sum_recycled vector by 3.
3. Try to interpret each result

## 1.2. Index by character
We can specify the name of each value using the equal (=) when creating the vector. 
```{r}
math_grades <- c("Exams"=92L, "Homework"=87L, "Projects"=85L)
print(math_grades)
```
You can use the names() to assign the characters into the vector
```{r}
grade_names <-  c("Exams", "Homework", "Projects")
math_grades <- c(92L, 87L, 85L)
names(math_grades) <- grade_names
```

To access/replace the value, we can use [] to access through position number or the name
```{r}
math_grades['Exams']
math_grades[1]
math_grades[c('Exams','Homework')]
math_grades[c(1,2)]

math_grades[c('Exams','Homework')] <-70L
print(math_grades)
```

### Exercise 2
```{r}
exam_grades <- c(92, 90, 84, 95, 77, 92, 85)
class_names <- c("math", "chemistry", "writing", "art", "history", "music", "physical_education")
```
1. Use the function names() to name elements in the vector exam_grades with the class names vector, class_names

# 2. Matrix
A matrix is a 2-dimensional data structure for storing values of SAME TYPE. It consists of rows and columns. One way to create a matrix is by restructuring a vector into a matrix. You can come up with two ways to do this using matrix().

```{r}
matrix_example <- matrix(1:8, nrow = 4, ncol = 2)
print(matrix_example)
```
You don't have to specficy both nrow and ncol.
```{r}
matrix_example2 <- matrix(1:8, nrow = 4)
print(matrix_example2)
matrix_example3 <- matrix(1:8, ncol = 2)
print(matrix_example3)
```
Note that you can fill each element by either rows or columns. Use byrow argument in matrix()
```{r}
matrix_example4 <- matrix(1:8, nrow = 4, ncol = 2, byrow = TRUE)
print(matrix_example4)
matrix_example5 <- matrix(1:8, nrow = 4, ncol = 2, byrow = FALSE)
print(matrix_example5)
```
### Exercise 3
```{r}
vector_1 <- c(1, 4, 5, 9) 
vector_2 <- 1:12 
vector_3 <- seq(from = 2.0, to = 5.50, by = 0.25) 
```
1. Create the matrix named matrix_1 matrix by restructuring vector_1 into two rows and columns.
2. Create the matrix named matrix_2 matrix by restructuring vector_2 into four rows and three columns.
3. Create the matrix named matrix_3 matrix by restructuring vector_3 into three rows and five columns.
- Set the parameter byrow to TRUE to restructure the vector by row.

## 2.1 Combine vectors
Vectors are stacked with elements aligned by position. rbind() and cbind() are used for stacking vectors. 
```{r}
world_rank <- c(1, 2, 3, 4, 5, 6)
quality_of_education <- c(1, 9, 3, 2, 7, 13)
influence <- c(1, 3, 2, 6, 12, 13)
broad_impact <- c(1, 4, 2, 13, 9, 12)
patents <- c(3, 10, 1, 48, 15, 4)

colum_stack <- cbind(world_rank,quality_of_education,influence,broad_impact,patents)
row_stack <- rbind(world_rank,quality_of_education,influence,broad_impact,patents)
print(colum_stack)
print(row_stack)
```
When we use the rbind() function to create a matrix from vectors with name attributes, the vector names are retained as row name attributes. However, our columns don't have any name attributes associated with them. Adding names to rows or columns of a matrix is similar to naming elements of a vector.
```{r}
print(colum_stack)
rownames(colum_stack)

school_names <- c("harvard", "stanford", "MIT", "cambridge", "oxford", "columbia")
rownames(colum_stack) <- school_names
print(colum_stack)
```
As with vectors, if we call the rownames() or colnames() functions in a matrix without assigned names, the function will return the names of rows or columns, or NULL if there is no name.

## 2.2 Index matrix
How to locate the value in matrix? The syntax for indexing matrices is similar to that for indexing vectors. We can extract values from the matrix by specifying their rows and/or columns in square brackets ([]) separated by comma(,).
```{r}
print(colum_stack)
```
How to access quality of education rank for Oxford university?
```{r}
colum_stack[5,2]
colum_stack['oxford','quality_of_education']
```
See the following
```{r}
colum_stack[5,]
colum_stack['oxford',]
colum_stack[,2]
colum_stack[,'quality_of_education']
```

### Exercise 4
```{r}
university_matrix <- colum_stack
```
1. Write code to get the value in the columbia row and influence column of university_matrix.
2. Write code to select elements in the cambridge row of university_matrix.
3. Write code to select elements in the world_rank column of university_matrix.

## 2.3 Slicing
Now we can apply what we have learned.

```{r}
print(university_matrix)
```
To return the values in the first three rows (harvard, stanford, and MIT) and the second column (quality_of_education) of university_matrix, we can write:
```{r}
university_matrix[1:3, 2]
university_matrix[c("harvard", "stanford", "MIT"), "quality_of_education"]
```
Try to explain the following code:
```{r}
usa_universities <- c(TRUE, TRUE, TRUE, FALSE, FALSE, TRUE) # e.g. the first TRUE means Harvard is an American university.
university_matrix[usa_universities, ] # Notice the absence of the column index
```
Guess the following code:
```{r}
usa_universities <- c(TRUE, FALSE)
university_matrix[usa_universities, ] 
```

### Exercise 5
1. Write code to index the values in the harvard, MIT, and cambridge rows and the world_rank and broad_impact) columns of university_matrix.
2. Write code to index the rankings for the USA's universities (i.e., Harvard, Stanford, MIT, and Columbia) in the world world_rank and patents categories.

## 2.4 Replacement
```{r}
print(university_matrix)
university_matrix[3, 2] <- 2 # modify the element at the third row and second column
print(university_matrix)
university_matrix[usa_universities, 3:5] <- 1
print(university_matrix)
```

## 2.5 Appending more information
cbind(): apend new column

rbind(): append new row
```{r}
tuition <- c(43280, 45000, 45016, 49350, 28450, 55161)
cbind(university_matrix, tuition)
```

### Exercise 6
Add the elements in california as a new row of university_matrix.
```{r}
california <- c(7, 5, 4, 7, 29, 43176)
```

## 2.6 Delete 
We can remove values from the matrix by specifying their row and column preceded by a minus symbol (-) in square brackets ([]). To remove the second row of complete_university_matrix, we can write:
```{r}
complete_university_matrix <- rbind(university_matrix, california)
print(complete_university_matrix)
complete_university_matrix[-2,]
```
Guess:
```{r}
usa_universities <- c(TRUE, TRUE, TRUE, FALSE, FALSE, TRUE, TRUE)
complete_university_matrix[usa_universities, c(-2, -3)]
```

To remove the second, third and fourth columns from complete_university_matrix where the tuition is greater than or equal to 45000, we can write:
```{r}
complete_university_matrix[tuition >= 45000, c(-2, -3, -4)]
```

### Exercise 7
1. Write code to remove the first column from complete_university_matrix.
2. Write code to remove the second and fifth columns where the broad_impact is less than or equal to 10, broad_impact <= 10.

## 2.7 Operations
Consider the following:
```{r}
sum(complete_university_matrix["harvard", -6])
sum(complete_university_matrix["stanford", -6])
sum(complete_university_matrix["MIT", -6])
```
However, we're learning to use R with larger datasets. Instead of calculating each university score, we can use the function rowSums() to perform these calculations. Hence, it won't be necessary to specify the row name.
```{r}
rowSums(complete_university_matrix[, -6])
```
We can also use the function colSums() to compute an aggregate ranking for each category summing rankings for every university.
```{r}
colSums(complete_university_matrix[, -6])
```
In R, there are the functions mean(), rowMeans(), and colMeans() to calculate the average scores, which have a similar syntax as the functions sum(), rowSums(), and colSums(), we have learned.

### Exercise 8
1. Compute an aggregate ranking for the University of Harvard by averaging rankings for every category using the function mean().
2. Compute an aggregate ranking for each university averaging rankings for every category using the function rowMeans().
3. Compute an aggregate ranking for each category averaging rankings for every university using the function colMeans().

## 2.8 One more thing
Let's analyze the university ranking data to decide on a university to attend. To answer this question, let's develop a score for each university that is an aggregate of the rankings and tuition information available in university_matrix. We'll rank the universities for each category we have data for: world_rank, quality_of_education, influence, broad_impact, patents, and tuition.

We use the rank() function, along with an expression that indexes the matrix, to specify the category we want to rank the universities by. The rank() function returns a vector of numeric values. For example, to rank universities by quality_of_education, we'd write:

```{r}
rank(university_matrix[,"quality_of_education"])
```
Consider:
```{r}
world_rank_rank <- rank(university_matrix[,"world_rank"])
quality_of_education_rank <- rank(university_matrix[,"quality_of_education"])
influence_rank <- rank(university_matrix[,"influence"])
broad_impact_rank <- rank(university_matrix[,"broad_impact"])
patents_rank <- rank(university_matrix[,"patents"])

ranks_university_matrix <- rbind(world_rank_rank, quality_of_education_rank, influence_rank, broad_impact_rank, patents_rank)
print(ranks_university_matrix)
colSums(ranks_university_matrix)
```

### Exercise 9
1. Calculate the average of each university's rankings in ranks_university_matrix using the colMeans() function.
