---
title: "Lecture5"
author: "Yeabin Moon"
date: "2022-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction to Introduction to R

Here are a few topics we will cover in this lesson:

- The basics of R syntax
- Introduction to data types in R
- How to perform calculations using comparison operators


## 1.1 Console and script
In the script or code editor, you can write the instructions. 
```{r}
1223189283+2
```
Run this commend either in the script or console.

What is [1] in front of the result? Try this:
```{r}
1:100
```

## 1.2 Data type
In R, there are five basic data types: Logical, Integer, Numeric, Character, and Complex.

Logical data type consists of TRUE and FALSE
```{r}
TRUE == T
TRUE == 1
```
```
Try: TRUE == True
```
R is case-sensitive!

There are two types of numbers in R: Integer and Numeric. You can think that whole numbers and other numbers. Note that R recognizes any number as Numeric type. We need to put "L" after whole number to make R understand it as Integer
```{r}
100L
99999L
```

To represent character, we need either quotation marks " or apsstrophes '
```{r}
"Hi"
"Have you put enough effort?"
"'FALSE' is not logical literal any more"
"'1' is a character"
```

# 2. Arithmetic
You know how to do arithmetic. I want to show something more.

## 2.1 R knows the operator priority rules
```{r}
1+2*3
(1+2)*3
```

## 2.2 Data type
Recall that there are two types of number. Operations among values of the same data type yield the same data type
```{r}
result1 <- 123L + 123L * 2L
print(result1)
class(result1)
result2 <- 123 + 123 * 2
print(result2)
class(result2)
```
Note that data types has a descending order: Numeric, Integer, Logical. Operations among different data type yield the highest data type 
```{r}
result3 <- 123L + 123L * 2
print(result3)
class(result3)
```

## 2.3 Variables
We save a result into a variable in many purposes. The variable is nothing but a position in your machine in which your computer saves the result. Each programming language has a rule on variable naming. Here is the R:

1. Name consists of letters, numbers, ., _ (no comma)

2. The first letter should be a letter or .
  - You cannot start with a number
  - You can start with . but then number cannot be followed
```
Try: .a<-2
Try: .2<-2
Try: 1998aa<-2
```
3. Special cahracters are not allowed (spaces, operators,", parentheses)

4. Don't use a keyword already used in R

```{r}
here_are_some_examples_including_spaces_variables <- 10
```

## 2.4 Relational Operators
Relational operators are used to compare between values. So it returns logical values either TRUE or FALSE.

- Less than (<)
- Greater than (>)
- Less than or equal to (<=)
- Greater than or equal to (>=)
- Equal to (==)
- Not equal to (!=)

```{r}
223L < 4000L
!TRUE == FALSE
```

Consider this:
```
11 - 23123 > 11 + 111111111
```
Note that all arithmetic expression are evaluated before logical expressions.

## 2.5 Logical operators
- Logical NOT (!)
- Logical AND (&&)
- Logical OR (||)
Consider:
```
TRUE && FALSE
TRUE || FALSE
!TRUE
```
<b>Try to think about how to code $-1< x< 2$. </b>

# 3. Data basics
## 3.1 Packages
R contains a lot of useful toolbox to perform data work. These tools are collection of functions which are stored in packages. 

To install an R package, we use the function <b>install.packages()</b>.
```
install.packages("readr")
```
This command instruct R to install the packages from Comprehensive R Archive Network (CRAN) repository. Note that we need to indicate the package name as a character string. After installing it, we need to <em>load</em> it on the console. To laad the packages,
```{r}
library(readr)
```

## 3.2 Load your data
The readr package is part of the tidyverse collection of packages, which is mainly used to import a dataset in R. The readr package contains a read_csv() function importing data in CSV format into R. It stores data as <em>dataframe</em>.

```{r}
library(tidyverse)
url <- "https://yeabinmoonecon.github.io/ibs_course/BUS240/homework/homework1/monster_jobs_clean.csv"
df <- read_csv(url)
```

What if you want to use data in your machine? There are two ways. First, you can type the exact location of your data. 
```
df_location <- "/Users/yeabinmoon/Downloads/monster_jobs_clean.csv"
df <- read_csv(url)
```
Second, set the working directory in which your data is located.
```
setwd("/Users/yeabinmoon/Downloads/")
df <- read_csv("monster_jobs_clean.csv")
```

## 3.3 Characterize dataset
```{r}
head(df,7)
tail(df,7)
glimpse(df)

n_cols <- ncol(df)
n_rows <- nrow(df)
names <- colnames(df)
```

## 3.4 Plot
```{r}
library(ggplot2)
```
The qplot() function of ggplot2 package allows representing graphs quickly.
```{r}
p <- qplot(x = job_id, y = salary_min, data = df)
p
```

```{r}
p <- qplot(x = job_id, y = salary_min, color = job_type ,data = df)
p
```

# 4. Data Structures 
## 4.1 Vector
We will explore four data structures:

- Vector: one-dimensional structure for storing values of SAME TYPE.
- Matrix: two-dimensional structure for storing values of SAME TYPE.
- Lists: multi-dimensional stucture for storing values of ANY DATA TYPE/OBJECT.
- Dataframe: two-dimensional structure for storing values of ANY DATA TYPE/OBJECT.


## 4.2 Sequence
A simple way to create vectors is to generate a sequence of values. 

1. Use : to create a range of values
```{r}
sample_seq1 <- 1:10
sample_seq1
```
2. Use seq() function
```{r}
sample_seq2 <- seq(from = 2, to = 20, by = 2)
sample_seq2
```
3. Use rep() function to have a repeated value
```{r}
sample_seq3 <- rep(5, times = 10)
sample_seq3
```

### Exercise
1. Create a vector that contains the value 4 repeated ten times using the rep() function.
2. Create a range of values from 1 to 12.
3. Create a sequence of values from 2.0 to 5.50 by 0.25.

## 4.3 Create a vector
A vector is created thanks to the function c() (combine, concatenate, collect,...). 

Note that there are four data types: name it!

```{r}
c(T,F,T)
c(12L,3L,2323L)
c(12,3.4,232/2)
c("Hi","are","you still there?")
```
Without c() R throws a syntax error.

## 4.4. Index

```{r}
BUS211_grade <- c(100L,90L,10L,88L)
BUS240_grade <- c(88,23,44,99)
good_class <- c(TRUE, F,F,F)
```
R is a 1-indexed programming language, which means that the first element in a vector is assigned a position of one. That is, to access the first grade in BUS211_grade
```{r}
BUS211_grade[1]
```
We can extract values from the vector by specifying their position in square brackets.

To access the first two grades in BUS240_grade
```{r}
BUS240_grade[1:2]
BUS240_grade[c(1,2)]
```

### Exercise
```{r}
math_grades <- c(92L, 87L, 85L)
exam_grades <- c(92, 90, 84, 95, 77, 92, 85)
```
1. Write code to index the first element of the vector, math_grades
2. Write code to select elements in the second, third, and sixth position of the exam_grades vector
3. Write code to select the third through seventh elements of the exam_grades vector

To access the first two grades in BUS240_grade
```{r}
BUS240_grade[1:2]
BUS240_grade[c(T,T,F,F)]
```
Explain the last code.

## 4.5 Replace
Modify the second element in BUS240_grade
```{r}
BUS240_grade[2] <- 77
print(BUS240_grade)
```

Predict the result of the codes:
```
BUS240_grade[c(T,T,F,F)] <- 89
print(BUS240_grade)
```
### Exercise
```{r}
math_grades <- c(92L, 87L, 85L)
exam_grades <- c(92, 90, 84, 95, 77, 92, 85)
is_stem_classes <- c(TRUE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE)
```
1. Replace the second element in math_grades by 80L
2. Replace the STEM exam grades with 99

## 4.6 Combine vectors
```{r}
BUS211_grade <- c(100L,90L,10L,88L)
extra_BUS211_grade <- c(99L, 79L)
new_BUS211_grade <- c(BUS211_grade, extra_BUS211_grade)
print(new_BUS211_grade)
```

## 4.7 Remove elements
```{r}
BUS211_grade <- c(100L,90L,10L,88L)
BUS240_grade <- c(88,23,44,99)
```
We can remove values from the vector by specifying their position preceded by a minus symbol (-) in square brackets
```{r}
BUS211_grade[-4]
```
But see:
```{r}
BUS211_grade
```
What's happening here?

You can use a vector to specify which elements you want to remove
```{r}
BUS240_grade[c(-1,-4)]
```
Make sure that you need to assign it to an object. Otherwise, none of the operations are maintained.

### Exercise
```{r}
math_grades <- c(92L, 87L, 85L)
exam_grades <- c(92, 90, 84, 95, 77, 92, 85)
```
1. Write code to remove the first element of the vector, math_grades.
2. Write code to remove exam grades less than 90, i.e., the element at the third, fifth and seventh positions of the exam_grades vector.